---
bibliography: "references.bib"
title: "tydygraphs"
output: 
  html_document:
    toc: true
    toc_float: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(knitcitations)
library(tidyverse)
cite_options(citation_format = "pandoc")
```
_(tie-dye graphs)_

Long time series datasets can be difficult to plot and examine using static
figures. Interactive plots using the
[dygraphs](https://rstudio.github.io/dygraphs/) package are a great tool but are
currently not easily implemented for tidy data. This package provides methods
for `dygraph()` to make it easier to plot tidy time series. Currently there are
methods for dataframes and tibbles, grouped tibbles, and tsibbles. 

## Installation

Currently tydygraphs is only available on GitHub. You can install the current
development version using [remotes](https://remotes.r-lib.org/) or by
downloading and building from source:

``` r
# Install using remotes
remotes::install_github("jpshanno/tydygraphs")

# or from source
tydygraphs_source <- file.path(tempdir(), "tydygraphs-master.zip")
download.file("https://github.com/jpshanno/tydygraphs/archive/master.zip",
              tydygraphs_source)
unzip(tydygraphs_source,
      exdir = dirname(tydygraphs_source))
install.packages(sub(".zip$", "", tydygraphs_source), 
                 repos = NULL,
                 type = "source")

```
## Use

### A Simple Example

The dataset `great_lakes_hydro` included in tydygraphs demonstrates the problems
with examing long-term time series datasets and the benefits of using dygraphs.
The dataset consists of monthly water levels, precipitation, evaporation,
discharge from, and runoff into Lake Superior and the combined Lakes Michigan
and Huron. The data were downloaded from [NOAA's Great Lakes Dashboard
Project](https://www.glerl.noaa.gov//data/dashboard/portal.html) `r citep("10.1016/j.envsoft.2015.12.005")`

```{r show-data}
library(tydygraphs)
great_lakes_hydro
```

Let's start simple and show how the we can use the function for a single time
series by filtering and supplying just a single variable to `dygraph`. The first
column representing a time series will be used if nothing is supplied for `time`.

```{r example}
great_lakes_hydro %>% 
  filter(lake == "Superior") %>% 
  dygraph(water_level_m)
```

### Adding More Series
And if you noticed in the data, all of the other variables are expressed in
millimeters, but because tydygraphs is built on top of dygraphs we can easily
add two a second y-axis to look at precipitation in the same plot. This plot
also makes it pretty clear why a tool like dygraphs is great for long time
series. When the entire dataset is displayed it is nearly useless for
exploratory analysis or diagnostics, but now we have the ability to zoom in on
periods of interest.

```{r example-2}
two_variables <- 
  great_lakes_hydro %>% 
  filter(lake == "Superior") %>% 
  dygraph(precip_mm,
          water_level_m) %>% 
  dySeries("precip_mm", 
           axis = 'y2')

two_variables
```

### Making it Look Nice

If we wanted to use this plot in something like a Shiny application or a blog
post we could build even more using the dygraphs customizations. The chunk below
makes precipitation into a step plot, adds a border around the water level
series, and starts the plot zoomed in on a period of interest.

There are lots of ways to customize dygraphs and I highly recommend looking at
the documentation for [dygraphs](https://rstudio.github.io/dygraphs/).

```{r example-3}
# Start with the plot from above
two_variables %>% 
  # Modify precipitation to make a filled step plot with no border by adjusting
  # the strokePattern
  dySeries("precip_mm",
           axis = "y2",
           stepPlot = TRUE,
           fillGraph = TRUE,
           color = "dodgerblue",
           strokePattern = c(0,1)) %>%
  # Modify water level to add a white border around the line to ensure it stands
  # out from precipitation
  dySeries("water_level_m",
           strokeBorderWidth = 2,
           color = "darkblue") %>%  
  dyRangeSelector(dateWindow = c("2000-01-01", "2010-01-01")) %>% 
  dyLegend(labelsSeparateLines = TRUE)
```

### Working with Grouped Data

And the final example is showing how we can plot our tidy data by passing a 
grouped tibble to `dygraph()`.

```{r example-4}
great_lakes_hydro %>% 
  group_by(lake) %>% 
  dygraph(water_level_m)
```



## Future Plans
- Add unit tests

### References
```{r message=FALSE, include=FALSE}
write.bibtex(file="references.bib")
RefManageR::PrintBibliography(read.bibtex(file = "references.bib"))
```
